---
title: "Assignment4"
author: "Tabea"
date: "16 November 2018"
output: 
  html_document: 
    keep_md: yes
---
#Data visualization with Base R
##Chapter 1
```{r}
Animals <-read.csv("C:/Users/Tabea/Downloads/Animals.csv", row.names = 1, sep = ";", header = TRUE)
```
###- 2 basic types of visualizations: 
### -> exploratory visualization helps understanding data
### -> explanatory visualization helps sharing with others

### -> exploratory: plot(dataset)
```{r}
plot(Animals)
```

### -> explanatory: plot(dataset$col1, dataset$col2, xlab = "Colname1", ylab = "Colname2")
```{r}
plot(Animals$Body_length, Animals$Body_weight)
```

###- factor varibles give barplot
### -> barplot()
```{r}
barplot(Animals$Age, horiz = TRUE)
```

###- title of plot (within function): main = "title"
```{r}
barplot(Animals$Age, horiz = TRUE, main="Animal Age")
```
###- sunflowerplot(dataset$col1, dataset$col2)
```{r}
sunflowerplot(Animals$Body_length, Animals$Body_weight)
```

###- boxplot(var1 ~ var2, data = dataset)
### -> how var1 varies across var2
```{r}
boxplot(Body_length~Body_weight, data=Animals)
```

###- mosaicplot(var1 ~ var2, data = dataset)
```{r}
mosaicplot(Age~Body_weight, data=Animals)
```

###- add features to a plot
### -> points() add points
### -> lines() add lines, usually curved
### -> text() add labels
### -> pch = shape of scatters
### -> col = colour
### -> abline(a = intercept, b = slope, lty = linetype, lwd = differing width)
```{r}
plot(Animals$Body_length, Animals$Body_weight)
points(Animals$Body_length, Animals$Age, pch=5, col= "red")
abline(a=0, b=10, lty=2, lwd= 3)
```

###- plot multiple plots on single pane: par(mfrow = c(#rows,   #cols)) 
### -> use par() to set graphical parameters

###- add title later (directly after plot()): title("title")
```{r}
plot(Animals$Body_length, Animals$Body_weight)
title("Weight vs Length")
```
###- log = "xy"
```{r}
plot(Animals$Body_length, Animals$Body_weight, log="y")
```
###- pie()  -> pie chart
### -> not good to use -> bar charts better
```{r}
pie(Animals$Age)
```
###- las = 2 makes labels perpendicular to axes
###- cex.names = 0.5 makes name labels half the default size

##Chapter 2

###- Histograms used for gaussian distributions of numerical values
### -> hist() -> histogram
### -> truehist() -> normalized histogram
```{r}
hist(Animals$Body_length)
```

###- which(data == "search term")

###- add density curve to histogram
### -> lines(density(dataset))

###- qqPlot()  -> quantile-quantile plot; only 1 argument
```{r}
qqplot(Animals$Body_weight, Animals$Body_length)
```
###- plot(var1 ~ var2, data = dataset )  -> var1=y, var2=x

###- optional boxplot parameters
### -> varwidth: variable-width boxplots with different sizes of data subsets
### -> log: log-transformed y-values
### -> las: more readable axis labels

###- mosaicplot ~ scatterplot for categorial variables
### -> mosaicplot(var1 ~ var2, data = dataset)

###- bagplot (aplpack package): 2dimensional boxplot
### -> cex = # : adjust size of point

###- correlation plot: correlation coefficient between each pair of variables as elipse (slope and roundness important)
### -> cor() to compute the correlation matrix with the correlation coefficients
### -> corrplot(correlationMatrix, method = "ellipse") 

###- decision trees (rpart package)
### -> rpart(var1 ~ var2 + var3 (or just "."), data = dataset), then plot()

##chapter 3

###par() -> set graphical parameters globally
### -> type values: "l"=lines, "p"=points, "o"=overlaid, "s"=steps, "n"=blank canvas (points(), lines(),... to add data)

###- Gaussian density: 
### -> dnorm(dataset, mean = __, sd = __)

###- linear regression model to predict y from x
### -> lm(y ~ x, data = df)

###- text(x,y,labels,adj,srt,font, cex, col)
### -> adj = 0.5 centered text, 0 left-& 1 right-justified
### -> srt = angle of text
### -> cex = size of text

###- legend("position", pch = c(1,2), legend = c("text1", "text2"))
```{r}
hist(Animals$Body_length)
legend(1, pch = 1, legend="Body Length")
```
###- axis()


###- supsmu(x,y,bass) adds two trend lines to scatterplot-> one using  default parameters and other increased smoothness(regulated by bass)

##Chapter 4

###- matplot(data1, dataset[,c("data2","data3")]) 
### -> multiple scatterplots on one set of axes

###- wordcloud(words = names(___), freq = as.numeric(___),           scale = ___, min.freq = ___)
### -> wordcloud package 

###- multiple plots -> set up array with par(mfrow=c(x,y)), then create plots

###- lauyout(matrix) uses matrix to define plot array
###- layout.show(n=number of plots)

##Chapter 5

###-symbols(data1, data2, circles=data3, inches=vary radius) -> show relation between >= 3 variables
```{r}
symbols(Animals$Body_length, Animals$Body_weight, circles = Animals$Age, inches = 0.2)
```

###- png(dataset.png) -> create png file
###- dev.off() -> exit current environment

###- suppressPackageStartupMessages(library()) -> avoid unncessary output in console

## Other packages for plot creation
###- tabplot -> tableplot()
###- lattice -> xyplot()
###- ggplot2 -> ggplot(dataset,aes(x=var1,y=var2))

